Ultrafast Electron Microscopes: Design Criteria, Electron Sources, and Column Modeling

### 2013-04-08 inc/photocathode/este/este.tex
moved more mentions of non-este m* from other sections to future

% the content of this section is a modification of our APL paper on ESTE
\section{Excited State Thermionic Emission}

@@

The dependence of $\Delta p_{\smallT}$ on $m^*$ is readily explained through a consideration of energy and momentum conservation in transmission across a boundary with a potential step associated with the effective photoelectric work function $\Phi$.
Inside the material, an electron with an energy $E$ greater than $\Phi$ has a maximum momentum of $\sqrt{2 m^* (E-\Phi) }$ parallel to the boundary if it is to be emitted.
As this momentum component is conserved in emission from the material surface into the vacuum, the rms transverse momentum $\Delta p_{\smallT}$ of the electron source clearly must scale with $\sqrt{m^*}$ as observed in the experiment.
This is, of course, the \textit{sine qua non} of angle-resolved photoemission studies \cite{himpsel_angle-resolved_1983} where determination of the electron emission angle and energy allows the effective mass parallel to the sample surface to be determined.
We note that the electron effective mass also causes the narrow emission cone reported for $p$-type GaAs(100) negative electron affinity photocathodes \cite{liu_narrow_2005}, where surface cesiation lowers the vacuum level below the $\Gamma_6$-valley minimum ($m^* = 0.067m$)  to allow for direct emission of electrons excited into the conduction band.
For direct single-photon photoemission, we would therefore also expect to require that the expression describing the rms transverse momentum to be written as $\Delta p_{\smallT} = \sqrt{m^* ( \hbar \omega - \Phi ) / 3 }$.
Efforts are currently underway to determine if this is indeed the case using a variety of metal photocathodes and the $\sim$4ps, 261nm laser radiation source.

>>>>

% the content of this section is a modification of our APL paper on ESTE
\section{Excited State Thermionic Emission} \label{sec:este}

@@

The dependence of $\Delta p_{\smallT}$ on $m^*$ is readily explained through a consideration of energy and momentum conservation in transmission across a boundary with a potential step associated with the effective photoelectric work function $\Phi$.
Inside the material, an electron with an energy $E$ greater than $\Phi$ has a maximum momentum of $\sqrt{2 m^* (E-\Phi) }$ parallel to the boundary if it is to be emitted.
As this momentum component is conserved in emission from the material surface into the vacuum, the rms transverse momentum $\Delta p_{\smallT}$ of the electron source clearly must scale with $\sqrt{m^*}$ as observed in the experiment.
This is, of course, the \textit{sine qua non} of angle-resolved photoemission studies \cite{himpsel_angle-resolved_1983} where determination of the electron emission angle and energy allows the effective mass parallel to the sample surface to be determined.
We note that the electron effective mass also causes the narrow emission cone reported for $p$-type GaAs(100) negative electron affinity photocathodes \cite{liu_narrow_2005}, where surface cesiation lowers the vacuum level below the $\Gamma_6$-valley minimum ($m^* = 0.067m$)  to allow for direct emission of electrons excited into the conduction band.

### 2013-04-08 inc/photocathode/future.tex
moved more mentions of non-este m* from other sections to future

>>>>

The philosophical argument presented in Section \ref{sec:este} for using the effective mass $m^*$ in the expressions for the transverse momentum variance $\Delta p_{\smallT}$ simply follows from conservation of energy and momentum at the barrier and actually does not depend on the ``excited state'' nature nor the thermionic final emission process.
The earlier experiment employing two-photon assisted thermionic emission (2PTE) from gold (Section \ref{sec:two_photon_thermionic}), the close agreement between the data and the simulation is supported by the fact that the effective mass of a free electron in Au is equal to the rest mass $m$ \cite{johnson_optical_1972}.

For direct single-photon photoemission, we would therefore also expect to require that the expression describing the rms transverse momentum to be written as $\Delta p_{\smallT} = \sqrt{m^* ( \hbar \omega - \Phi ) / 3 }$.
Efforts are currently underway to determine if this is indeed the case using a variety of metal photocathodes and the $\sim$4ps, 261nm laser radiation source.
In an upcoming publication, we will show that the dependence of $\Delta p_{\smallT}$ on the electron effective mass may be extended to simple planar metal photocathodes --- thus potentially allowing for the future development or discovery of a robust, high brightness photocathode material.
The simulation is in good agreement with our direct measurements of $\Delta p_{\smallT}$ for Bi(111) and polycrystalline Ag, Mo, Ta, Be, and W photocathodes.
We will also discuss the effect of laser-induced heatingon $\Delta p_{\smallT}$ of the electron distribution in the photocathode during photoemission, which may limit the attainable pulsed electron source brightness at the desired high peak pulse current densities.
